\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-09 doc/publications/plos16/plos16.tex
Few fix ups of previous commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
In this section we use \cursedforest\ to identify genomic locations that are associated with a specific trait. To do
this, we perform feature selection over 7.2 million genomic variants to identify the locations associated with Bone
Mineral Density (BMD)~\cite{Duncan.2011}. We build a model on the BMD dataset (see datasets) using a binary high or low
BMD value as the label.  From this model, we output a list of the top 10,000 contributing variants, ranked by their Gini
impurity. The variant with the highest Gini impurity, and hence importance, is ranked 1, and the variant with the lowest
Gini impurity is ranked 10,000.

To test whether the associated features have biological meaning, we investigate whether the known BMD genes investigated
by Duncan {\it et al.}~\cite{Duncan.2011} are enriched amongst the highly ranked genes. We therefore label all variants
that fall within 100 kb of these 32 BMD genes and observe that they are ranked significantly higher than other locations
in the genome (Mann-Whitney U, $p=1.3\times10^{-07}$), supporting our hypothesis that biologically relevant features are
contributing more information to our classification model.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
In this section we use \cursedforest\ to identify genomic locations that are associated with a specific trait. To do
this, we perform feature selection on over 7.2 million genomic variants to identify the locations associated with Bone
Mineral Density (BMD)~\cite{Duncan.2011}. We build a model on the BMD dataset (see datasets) using a binary high or low
BMD value as the label.  From this model, we output a list of the top 10,000 contributing variants, ranked by their Gini
impurity. The variant with the highest Gini impurity, and hence importance, is ranked 1, and the variant with the lowest
Gini impurity is ranked 10,000.

To test whether the associated features have biological meaning, we investigate whether variants in or near
BMD-associated genes are enriched toward the top of our ranked variant list. To form this list of "known genes", we start
with the 26 known BMD genes listed by Duncan {\it et al.}~\cite{Duncan.2011}. We also include the 6 genes that they find
with a new BMD association. We then add to this list the genes that Estrada {\it et al.} found with a new BMD association.
In total, this gives us a list of 100 genes.

We label variants from our ranked list that fall within 100 kb of these 100 genes and observe that they are ranked
significantly higher than variants from other locations in the genome (Mann-Whitney U, $p=1.3\times10^{-07}$),
supporting our hypothesis that biologically relevant features are contributing more information to our classification model.
